---
title: "Calc_metrics_prelimAnalysis"
author: "Alexandra Alexiev"
date: "4/21/2021"
output: html_document
---
### Packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


### Load data

Previously, we fit growth curves to three data sets:

1) CFS as
Cell-free supernatant experiment to test additive vs. synergistic effects of probiotics vs. cocktails on Bd growth

2) CFS fac
Cell-free supernatant experiment to test facilitation of cocktail microbes against Bd growth

3) Bact fac
Bacterial growth only (no Bd) test to see if cocktail microbes increase each other's growth/growth rate more than when grown alone

```{r load data}
fit_cfs_as <- readRDS("output_files/fit_cfs_as.rds")
fit_cfs_fac <- readRDS("output_files/fit_cfs_fac.rds")
fit_bact_fac <- readRDS("output_files/fit_bact_fac.rds")

# assign objects to data frames of relevance
cfs_as_fitEst <- fit_cfs_as$`Pooled fit`
cfs_fac_fitEst <- fit_cfs_fac$`Pooled fit`
bact_fac_fitEst <- fit_bact_fac$`Pooled fit`
```


### Calculate metrics

Put fit estimates into the following equation: 

N1 = K/ (1 + (K-N0)/N0 * e^(-rt) )

* **K**: Carrying capacity
* **r**: Growth rate (h-1)
* **t**: Generation time
* **sigma**: Variance in data around the fit model
* **N0**: Estimate starting population
* **t_mid**: 1/2kmax in a MM model
* **N1**: Estimate end populations (we want population at 100 hrs)

```{r calc metrics}
cfs_as_fitEst$N1 <- with(cfs_as_fitEst, k / (1 + (k - n0)/n0 * exp(-(r * 100))))
cfs_as_fitEst[is.na(cfs_as_fitEst)] <- 0 # replace NaNs with zero since they are 0 growth

cfs_fac_fitEst$N1 <- with(cfs_fac_fitEst, k / (1 + (k - n0)/n0 * exp(-(r * 100))))
cfs_fac_fitEst[is.na(cfs_fac_fitEst)] <- 0 # replace NaNs with zero since they are 0 growth

bact_fac_fitEst$N1 <- with(bact_fac_fitEst, k / (1 + (k - n0)/n0 * exp(-(r * 100))))
bact_fac_fitEst[is.na(bact_fac_fitEst)] <- 0 # replace NaNs with zero since they are 0 growth
```

```{r save r data files}
# commented out so I don't overwrite existing saved files unless I want to

# save.image("~/Documents/Projects/ProbioticCocktails/02_PreliminaryAnalysis/Cocktail_trials/01_preliminaryAnalysis/Calc_metrics_prelimAnalysis.RData")
# 
# saveRDS(cfs_as_fitEst, "cfs_as_fitEst.rds")
# saveRDS(cfs_fac_fitEst, "cfs_fac_fitEst.rds")
# saveRDS(bact_fac_fitEst, "bact_fac_fitEst.rds")
```

## Cross-check visually that N1 calculated matches the fit on the graphs

Notes:

- Live Bd in both CFS trials was pretty similar magnitude
- Most things inhibited, some less than others
- 54D did not inhibit, as expected since it shouldn't
- N1 calculated seems to match where the line of best fit lands on the graphs
